---
title: "PSY 8960 Final Project Shiny App Notebook"
author: "Isaac Bazian"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **Script Settings and Resources**

```{r script_settings_and_resources, message=FALSE}
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
library(tidyverse)
```

## **Online Resources**


## **Data Import and Cleaning**

```{r data_import}
#This code reads in the original data
finalproj_full_tbl <- readRDS("../data/finalproj_data.rds")

# This code takes the original data and trims it down into a 'skinny' format, including
# only the variables that will be used in the app - these are the 3 possible outcomes
# and the 4 possible filter categories. It should be noted that I changed Attrition from
# 'Yes' and 'No' to 1 and 0 instead, making it numeric so that it could be properly
# plotted on a histogram. I add text in the app under the plot when the user selects this
# outcome to ensure they understand what 1 and 0 means in that context.
finalproj_skinny_tbl <- finalproj_full_tbl %>% 
  select(MonthlyIncome, Attrition, JobSatisfaction, Department, EducationField, Gender, JobRole) %>% 
  mutate(Attrition = case_match(Attrition, "Yes" ~ 1, "No" ~ 0)) #Also makes the column numeric

 #This code takes the skinny data and stores it for use in the shiny app
saveRDS(finalproj_skinny_tbl, "../shiny/people_dashboard/finalshinydata.rds")
```



## **Justifications for Approach**

I chose to have users select which outcomes they wanted to see and what filtering
they wanted applied to the data using radio buttons. I personally think radio buttons
are quite intuitive, and they allow you to always see all available options.
A dropdown menu might have made things more compact, but I like being able to change
parameters with one click on a button, rather than clicking into a dropdown and then
clicking the new options. I used a sidebar layout because I thought it kept the buttons'
relations to the plot and table easy to intuit and see.

For the outcome, I have Monthly Income selected by default. This seemed like a reasonable
first thing a user would want to know, but it's ultimately arbitrary - any of the 3
outcomes could have been the default. Filters are set to 'All' by default, reflecting no
filtering. This way, the user starts with as much data as possible, which they can
then opt to cut down.

The app displays a histogram of the selected outcome with whatever data remains 
after filtering. This fulfilled the project requirement of a univariate visualization
in a way I thought was intuitive and informative. A table of means and standard deviations
of the selected outcome split by all possible groups of selected filtering terms
was also included to fulfill project requirements. Additionally, if the user selects
Turnover Status as the outcome of interest, I display text underneath the plot so that
they know a 0 means the employee did not turn over, and a 1 means the employee
did turn over.

Moving on to the server, I executed the logic in ways that made sense to my thought process.
Starting with the plot: The plot is ultimately going to be made in a pretty similar way,
it's just a question of what data makes it through the filtering. Therefore, after
reading in the skinny data, I make another copy of the data that will be successively filtered
as the user selects different filtering options. For any filter category where
the user leaves it at 'All' (or changes it back to 'All' later), no filtering is done
with that category. If they do select filtering in that category, whatever option
they choose is used to filter the data, such that it only keep rows in that filtered
column that exactly match the user's choice. Once the data goes through all of
its filtering, it can be fed into the plot.

The names of the outcomes users select do not exactly match the names of the variables
in the dataset, so I wrote logic to change the contents of an object to the relevant
variable name based on outcome the user chose. This also gets fed into the plot,
defining the x axis data and name, as well as the plot name.

If the user selects Turnover Status as their outcome of interest, there is logic to display
a text message under the plot explaining how to interpret 0s and 1s, as described above.

For displaying the table of means and standard deviations, the project 
requirements immediately made me think of tidyverse's group_by and summarise functions.
I knew it was only a matter of taking any filter variables that had an option 
other than 'All' selected, then grouping the original data by those variables and
summarising means and standard deviations. That's basically all that code does -
if a filter has any option but 'All' selected, the name of the variable is added to
a chacter vector, which is then fed into group_by (with some extra code to make
a character vector work) on the original data. The outcome selected is translated into
the variable name in the same way as it was for the plot, and fed into the pipeline to
get the mean and standard deviation. This results in a table of means and standard deviations
for the outcome selected, split by all combinations of filters applied.

With that all done, it's simply a matter of deploying the app to shinyapps and
putting the link to it at the top of this Rmarkdown.









